# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

#Unzip the file
```{r}
if (!file.exists("activity.csv")) {
  unzip("activity.zip", overwrite = TRUE)
}
```

```{r}
df  <-  read.csv('activity.csv',colClasses = c("numeric", "character", "integer"))
#df$date  <- strptime(df$date, "%Y-%m-%d")
df$date  <- as.Date(df$date, "%Y-%m-%d")
```

```{r}
colnames(df)
lapply(df,class)
```


## What is mean total number of steps taken per day?

aggregate data per day by removing the na values
```{r}
df1  <- aggregate(df$steps, by=list(date = df$date), FUN=sum,na.rm=TRUE)
```

Plot the histogram of number of steps taken per day
```{r}
hist(df1$x,breaks=30)
```

Mean value of number of steps taken per day. 
```{r}
mean(df1$x)  #9354.23
```

Median of number of steps taken per day
```{r}
median(df1$x) #10395
```

## What is the average daily activity pattern?

```{r}
df2  <- aggregate(df$steps, by=list(interval = df$interval), FUN=mean,na.rm=TRUE)
plot(df2$interval, df2$x, type="l", main="average daily pattern",
     xlab="inervals", ylab="numnber of steps",xlim=c(0, 2359))
```

Interval when the number of steps taken was maximum
```{r}
df2$interval[df2$x == max(df2$x)]
```
at 8:35 AM the avegare number of steps taken was maximum

## Imputing missing values

number of entried where steps data was not-known (NA)
```{r}
sum(is.na(df$steps))
```

Create a new dataset equal to original dataset but with missing values imputed
code adapted from [here](http://stackoverflow.com/questions/13114812/imputation-in-r)
```{r}
#install.packages("Hmisc")
library(Hmisc)
df3  <- df
df3$steps  <- with(df3, impute(steps, mean))
```

Plot the histogram of number of steps taken per day
```{r}
hist(df1$x,breaks=30)
```

Mean value of number of steps taken per day. 
```{r}
mean(df1$x) #9354.23
```

Median of number of steps taken per day
```{r}
median(df1$x) # 10395
```

## Are there differences in activity patterns between weekdays and weekends?


```{r}
df3_weekend  <- subset(df3,weekdays(df3$date) %in% c("Saturday","Sunday"))
df3_weekend$day  <- "weekend"
df3_weekday  <- subset(df3,!weekdays(df3$date) %in% c("Saturday","Sunday"))
df3_weekday$day  <- "weekday"
df4  <- rbind(df3_weekend, df3_weekday)
```


```{r}
df4$day  <- as.factor(df4$day)
class(df4$day)
```



```{r}
df5  <- aggregate(df4$steps, by=list(interval = df4$interval,day = df4$day), FUN=mean)
```


```{r}
library(ggplot2)
ggplot(data = df5, aes(interval, x, group=day)) +
  geom_line() +
  facet_grid(day ~. ) +
  xlab("Time Interval") +
  ylab("Mean Number of Steps") +
  ggtitle("Mean Number of Steps versus Time Interval")
```

